Supplementary Material for the article \textit{Quotient Normalized Maximum Likelihood Criterion for Learning Bayesian Network Structures}

### 2018-02-19 experiments.tex
Add acknowledgements

For each train/test sample, we ranked the predictive performance of
the models learned by four different structure learning criteria (rank
1 being the best and 4 the worst). Table~\ref{tbl:preds} features the
average rank for different data sets, the average being taken over
1000 different train test samples for each 9 sample sizes for each
data.  BIC's bias for simplicity makes it often win (written bold in
the table) with small sample sizes, but it often performs worst
(underlined) for the larger sample sizes (for the same reason), while
fNML seems to be good for large sample sizes. The striking feature
about the qNML is its robustness.  It is usually between BIC and fNML
for all the sample sizes making it a ``safe choice''. This can be
quantified if we further average the columns of
Table~\ref{tbl:preds}.  This yields the average ranks of $2.95, 2.57,
2.10$, and $2.37$, with standard deviations $0.49, 0.90, 0.76$, and
$0.43$.  While fNML achieves on average the best rank, the runner-up
qNML has the lowest standard deviation.

Figure~\ref{fig:bcnpmean} shows how fNML still sometimes behaves
strangely in terms of model complexity, here measured by the number of
parameters in the model. qNML, instead, appears to yield more
parsimonious models. To study the concern of fNML producing too
complex models for small sample sizes, we studied the number of
parameters in models produced by different model selection criteria
when using 10\% of each data set for structure learning.

@@

\begin{table}
  \caption{Average number of parameters in models 
    for different model selection criteria in 20 different data sets.}
\label{tbl:nofparams}
\begin{center}
\begin{tabular}{crrrrr}
%%    Data &     N &              BDeu &               BIC &           fNML &              qNML \\
       Data
    & N
    & \multicolumn{1}{p{0.7cm}}{\centering BDeu\\ }
    & \multicolumn{1}{p{0.8cm}}{\centering BIC\\ }
    & \multicolumn{1}{p{0.9cm}}{\centering fNML\\}
    & \multicolumn{1}{p{0.9cm}}{\centering qNML\\}\\
\midrule
    Iris &    15 &     \underline{37} &   \textbf{23} &                33 &               29 \\
 PostOpe &    18 &   \underline{1217} &   \textbf{19} &               397 &              146 \\
   Ecoli &    34 &    \underline{182} &   \textbf{31} &               162 &               77 \\
   Liver &    35 &                 45 &   \textbf{15} &    \underline{61} &               24 \\
    Wine &    36 &  \underline{16521} &   \textbf{70} &               807 &              205 \\
   Glass &    44 &   \underline{1677} &   \textbf{48} &               506 &               97 \\
 Thyroid &    44 &                 40 &   \textbf{23} &    \underline{66} &               28 \\
 HeartSt &    54 &  \underline{16861} &   \textbf{44} &              1110 &              256 \\
 BreastC &    58 &  \underline{25797} &   \textbf{49} &              3767 &              844 \\
 HeartHu &    60 &   \underline{1634} &   \textbf{43} &               792 &               90 \\
 HeartCl &    62 &  \underline{34381} &   \textbf{47} &              1433 &              404 \\
 BcWisco &    70 &   \underline{4630} &   \textbf{42} &               603 &               89 \\
 Diabete &    77 &                 39 &   \textbf{22} &   \underline{216} &               34 \\
 TicTacT &    96 &  \underline{13701} &   \textbf{25} &              1969 &              767 \\
 Balance &   126 &        \textbf{20} &            24 &                49 &  \underline{611} \\
   Yeast &   149 &                 71 &   \textbf{31} &   \underline{265} &               75 \\
 Abalone &   418 &                 91 &   \textbf{46} &   \underline{150} &               63 \\
 PageBlo &   548 &    \underline{703} &   \textbf{45} &               380 &               56 \\
 Shuttle &  5800 &                535 &   \textbf{99} &   \underline{717} &              130 \\
   Adult &  6513 &                699 &  \textbf{479} &  \underline{1555} &              945 \\
\end{tabular}
\end{center}
\end{table}

>>>>

For each train/test sample, we ranked the predictive performance of
the models learned by the four different scores (rank
1 being the best and 4 the worst). Table~\ref{tbl:preds} features the
average rank for different data sets, the average being taken over
1000 different train/test samples for each 9 sample sizes.
BIC's bias for simplicity makes it often win (written bold in
the table) with small sample sizes, but it performs worst
(underlined) for the larger sample sizes (for the same reason), while
fNML seems to be good for large sample sizes. The striking feature
about the qNML is its robustness.  It is usually between BIC and fNML
for all the sample sizes making it a ``safe choice''. This can be
quantified if we further average the columns of
Table~\ref{tbl:preds}, yielding the average ranks of $2.95, 2.57,
2.10$, and $2.37$, with standard deviations $0.49, 0.90, 0.76$, and
$0.43$.  While fNML achieves on average the best rank, the runner-up
qNML has the lowest standard deviation.

Figure~\ref{fig:bcnpmean} shows how fNML still sometimes behaves
strangely in terms of model complexity as measured by the number of
parameters in the model. qNML, instead, appears to yield more
parsimonious models. To study the concern of fNML producing too
complex models for small sample sizes, we studied the number of
parameters in models produced by different scores when using
10\% of each data set for structure learning.

@@

\begin{table}
  \caption{Average number of parameters in models 
    for different scores in 20 different data sets.}
\label{tbl:nofparams}
\begin{center}
\begin{tabular}{crrrrr}
%%    Data &     N &              BDeu &               BIC &           fNML &              qNML \\
       Data
    & N
    & \multicolumn{1}{p{0.7cm}}{\centering BDeu\\ }
    & \multicolumn{1}{p{0.8cm}}{\centering BIC\\ }
    & \multicolumn{1}{p{0.9cm}}{\centering fNML\\}
    & \multicolumn{1}{p{0.9cm}}{\centering qNML\\}\\
\midrule
    Iris &    15 &     \underline{37} &   \textbf{23} &                33 &               29 \\
 PostOpe &    18 &   \underline{1217} &   \textbf{19} &               397 &              146 \\
   Ecoli &    34 &    \underline{182} &   \textbf{31} &               162 &               77 \\
   Liver &    35 &                 45 &   \textbf{15} &    \underline{61} &               24 \\
    Wine &    36 &  \underline{16521} &   \textbf{70} &               807 &              205 \\
   Glass &    44 &   \underline{1677} &   \textbf{48} &               506 &               97 \\
 Thyroid &    44 &                 40 &   \textbf{23} &    \underline{66} &               28 \\
 HeartSt &    54 &  \underline{16861} &   \textbf{44} &              1110 &              256 \\
 BreastC &    58 &  \underline{25797} &   \textbf{49} &              3767 &              844 \\
 HeartHu &    60 &   \underline{1634} &   \textbf{43} &               792 &               90 \\
 HeartCl &    62 &  \underline{34381} &   \textbf{47} &              1433 &              404 \\
 BcWisco &    70 &   \underline{4630} &   \textbf{42} &               603 &               89 \\
 Diabete &    77 &                 39 &   \textbf{22} &   \underline{216} &               34 \\
 TicTacT &    96 &  \underline{13701} &   \textbf{25} &              1969 &              767 \\
 Balance &   126 &        \textbf{20} &            24 &                49 &  \underline{611} \\
   Yeast &   149 &                 71 &   \textbf{31} &   \underline{265} &               75 \\
 Abalone &   418 &                 91 &   \textbf{46} &   \underline{150} &               63 \\
 PageBlo &   548 &    \underline{703} &   \textbf{45} &               380 &               56 \\
 Shuttle &  5800 &                535 &   \textbf{99} &   \underline{717} &              130 \\
   Adult &  6513 &                699 &  \textbf{479} &  \underline{1555} &              945 \\
\end{tabular}
\end{center}
\end{table}
